Response to the Decision Letter of the Editor and Reviewers\\ \normalsize -- Wind veer and speed in turbulent Ekman flow part I: scaling analysis and velocity profile model --

### 2025-04-21 write1/REVISION1/responses.tex
continued reviewer response

\onehalfspacing

@@

\title{Response to Reviewers} 
\author{Cedrick Ansorge and Hauke Wurps }
\begin{document}
\maketitle

%-------------------------------------------------------------------------------------%

%######################################################################################%
% not all comments are included, only where action is needed.
%######################################################################################%

\section*{Response to Reviewer 1}

@@

@@

\revpoint{In line 47, friction velocity is denoted as $u_* = \sqrt{|\tau_{sfc}/\rho}$. However, in Eq. (1a) (below line 179), $\tau_{sfc}$ is defined as the kinematic shear stress. Please make them consistent. }{Manz thanks for spotting this inconsistency in nomenclature. We have aligned $\tau_\text{sfc}$ with the conventional definition of dynamic shear stress and now include the division bz density in Eq. (1b).}{\begin{subequations}\begin{equation}
    \underline{\tau}_\mathrm{sfc} = \left(\begin{array}{c} \tau_x\\ \tau_y\\\tau_z\end{array}\right) = - \rho\nu \left( \frac{\partial U}{\partial z} \hat{e}_x + \frac{\partial V}{\partial z} \hat{e}_y\right)
  \end{equation}
  and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and
  (ii) the coordinate system aligned with the free-atmosphere geostrophic wind labelled by an upper index $G$ as in $\underline{U}^G$.
  %
  We denote the square root of the modulus of surface shear, the surface friction, by
  \begin{equation}
    u_\star = \sqrt{ \frac{|| - \underline{\tau}_\mathrm{sfc} ||}{\rho} } 
    \label{eqn:ustar_def}
  \end{equation} and let $Z_\star=G/u_\star$;
\end{subequations}}

@@

\revpoint{Figure 4: Should the legend be $Z_* = 4 log(Re_D) -8$, instead of $u_\star* = ...$"?}{Indeed, there was a typo, please find the revised Figure below.}{
\includegraphics[width=\textwidth]{Fig4.eps}
}
\newpage
\section*{Response to Reviewer 2}
\revpoint{The authors perform direct numerical simulations of the Ekman layer at moderate Reynolds number, and use their simulations to examine the vertical mean velocity profiles in rotating, neutrally-stratified flows that serve as a proxy for the neutral atmospheric boundary layer. They compare their DNS results to a model that matches the Ekman model (in the outer layer) with the classical results for the inner layer (viscous layer plus log layer) in wall-bounded turbulent flows. Specifically, they examine the scaling of the spanwise velocity profile, and find that a matching procedure—combining both inner- and outer-scaling—is necessary. The article is generally clear and well-written, and addresses the important problem of how velocity profiles behave in the Ekman layer. Despite being an old problem, it has not been resolved, and there are many subtleties that are often neglected in the literature. I am overall supportive of this manuscript, but below I provide some specific comments which I believe will help improve the clarity of communication in the article. I am recommending that the manuscript is accepted in BLM, pending minor revisions. }{We appreciate your summary of our work and thank you for the time spent on our manuscript. Below, please find the detailed responses to your comments.}{}

@@

\newpage
\section*{Response to Reviewer 3 (Prof. Sukanta Basu)}

@@

\noindent We ask for your understanding that, as you already anticipated, a complete reconsideration of our results in the framework of your 2023 paper is not carried out at this point. We, however, would be happy to provide processed and unprocessed data if a colleague or potentially member of your working group is interested in carrying these ideas further.\\[1em]

@@

\revpoint{At the heart of this paper is Eq.~10a. Following Etling's work and others, this equation has been derived in Appendix A1. \newline 
  \[
    \frac{1}{G} \left(\begin{array}{c}U_\text{Ek} \\ V_\text{Ek} \end{array}\right) = \left(\begin{array}{c} 1 \\0 \end{array}\right) + e^{-z_\text{Ek}} \left[ a_\text{ek} \left(\begin{array}{c} -\cos z_\text{ek} \\ \sin z_\text{ek} \end{array}\right)+ b_\text{ek}\left(\begin{array}{c}\sin z_\text{ek} \\ \cos z_\text{ek}\end{array}\right)   \right]
  \] \newline  
  The problem with this equation is the assumption of constant eddy viscosity. As the authors showed in
  Figure 2c, the eddy viscosity profile from the DNS is far from constant. It has a clear maximum and a
  specific shape, which depends on Re.
  So, the authors took an empirical (engineering) approach on page 14 (line 403). They estimated a bulk
  eddy viscosity indirectly. This is a major weakness of this paper and could be avoided. Please see the
  following paper where we analytically derived the eddy viscosity profile from the Ekman equations.
  Since this is a relatively new paper, the authors might not have come across it. That is why I am revealing
  my identity.
  \begin{quotation}Basu, Sukanta, and Albert AM Holtslag. “A novel approach for deriving the stable
  boundary layer height and eddy viscosity profiles from the Ekman
  equations.” Boundary-Layer Meteorology 187.1 (2023): 105-115.
  \end{quotation}
  The authors might want to revise their derivations based on our analytical eddy viscosity profile. If that
  is difficult to do, I suggest highlighting the limitations of their current bulk approach. Also, emphasize
  the limitations of Eq. (10a) as it assumes constant eddy viscosity. }{}{}

>>>>

\renewcommand{\familydefault}{\sfdefault}
\onehalfspacing

@@

\newcommand{\explainformat}{\noindent In our response below, we repeat your original comment \textcolor{gray}{in gray font}. Next comes \textcolor{blue}{\textit{our response in blue font}}, and we close by the changes to the manuscript in regular font. Comments are separated by horizontal lines for readability.\\[1em]}

\title{Response to the Decision Letter of the Editor and Reviewers\\
\normalsize -- Wind veer and speed in turbulent Ekman flow part I: scaling analysis and velocity profile model -- }
\author{Cedrick Ansorge and Hauke Wurps }
\begin{document}
\maketitle
\tableofcontents 
\newpage 
%-------------------------------------------------------------------------------------%

%######################################################################################%
% not all comments are included, only where action is needed.
%######################################################################################%
\begin{flushright}{Berlin, \today}\\ \end{flushright}
\section{Response to the Editor} 
\noindent Dear Professor Anderson, \\[1em]

\noindent we thank you very much for the assessment of both parts of this combined manuscript and the commitment of BLM to consider this body of work as a whole. \\[1em]

\noindent Below, we detail our responses including changes incurred to the manuscript. We also provide an annotated manuscript (manuscript\_diff.pdf) to ease the assessment of changes and expedit the further processing of our manuscript. \\[1em]

\noindent As it happens, some changes to this part of the manuscript are required due to Reviewer comments on the second part of the paper. We detail these changes below -- before the responses to the individual reviewers. \\[1em]

\explainformat{}

\noindent Kind Regards, \\
\noindent Cedrick Ansorge and Hauke Wurps

\newpage 
\begin{flushright}{Berlin, \today}\\ \end{flushright}
\section{Response to Reviewer 1}

@@

\explainformat{}

@@

\revpoint{In line 47, friction velocity is denoted as $u_* = \sqrt{|\tau_{sfc}/\rho}$. However, in Eq. (1a) (below line 179), $\tau_{sfc}$ is defined as the kinematic shear stress. Please make them consistent. }{Many thanks for spotting this inconsistency in nomenclature. We have aligned $\tau_\text{sfc}$ with the conventional definition of dynamic shear stress and now include the division bz density in Eq. (1b).}{\begin{subequations}\begin{equation}
    \underline{\tau}_\mathrm{sfc} = \left(\begin{array}{c} \tau_x\\ \tau_y\\\tau_z\end{array}\right) = - \rho\nu \left( \frac{\partial U}{\partial z} \hat{e}_x + \frac{\partial V}{\partial z} \hat{e}_y\right)
  \end{equation}
  and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and
  (ii) the coordinate system aligned with the free-atmosphere geostrophic wind labelled by an upper index $G$ as in $\underline{U}^G$.
  %
  We denote the square root of the modulus of surface shear, the surface friction, by
  \begin{equation}
    u_\star = \sqrt{ \frac{|| - \underline{\tau}_\mathrm{sfc} ||}{\rho} } 
    \label{eqn:ustar_def}
  \end{equation} and let $Z_\star=G/u_\star$;
\end{subequations}}

@@

\revpoint{Figure 4: Should the legend be $Z_* = 4 log(Re_D) -8$, instead of $u_\star* = ...$"?}{Indeed, there was a typo, please find the revised Figure below.}{
\includegraphics[width=\textwidth]{Fig4.eps}
}
\newpage
\begin{flushright}{Berlin, \today}\\ \end{flushright}
\section{Response to Reviewer 2}
Dear Reviewer 2, \\[1em]

\explainformat{}

\noindent Cedrick Ansorge and Hauke Wurps  \\[0.5em]

\revpoint{The authors perform direct numerical simulations of the Ekman layer at moderate Reynolds number, and use their simulations to examine the vertical mean velocity profiles in rotating, neutrally-stratified flows that serve as a proxy for the neutral atmospheric boundary layer. They compare their DNS results to a model that matches the Ekman model (in the outer layer) with the classical results for the inner layer (viscous layer plus log layer) in wall-bounded turbulent flows. Specifically, they examine the scaling of the spanwise velocity profile, and find that a matching procedure—combining both inner- and outer-scaling—is necessary. The article is generally clear and well-written, and addresses the important problem of how velocity profiles behave in the Ekman layer. Despite being an old problem, it has not been resolved, and there are many subtleties that are often neglected in the literature. I am overall supportive of this manuscript, but below I provide some specific comments which I believe will help improve the clarity of communication in the article. I am recommending that the manuscript is accepted in BLM, pending minor revisions. }{We appreciate your summary of our work and thank you for the time spent on our manuscript. Below, please find the detailed responses to your comments.}{}

@@

\newpage
\section{Response to Reviewer 3 (Prof. Sukanta Basu)}

@@

\noindent We ask for your understanding that, as you already anticipated, a complete reconsideration of our results in the framework of your 2023 paper is not carried out at this point (the detailed reasoning is given below). We, however, would be happy to provide processed and unprocessed data if a colleague or potentially member of your working group is interested in carrying these ideas further.\\[1em]

@@

\revpoint{At the heart of this paper is Eq.~10a. Following Etling's work and others, this equation has been derived in Appendix A1. \newline 
  \[
    \frac{1}{G} \left(\begin{array}{c}U_\text{Ek} \\ V_\text{Ek} \end{array}\right) = \left(\begin{array}{c} 1 \\0 \end{array}\right) + e^{-z_\text{Ek}} \left[ a_\text{ek} \left(\begin{array}{c} -\cos z_\text{ek} \\ \sin z_\text{ek} \end{array}\right)+ b_\text{ek}\left(\begin{array}{c}\sin z_\text{ek} \\ \cos z_\text{ek}\end{array}\right)   \right]
  \] \newline  
  The problem with this equation is the assumption of constant eddy viscosity. As the authors showed in
  Figure 2c, the eddy viscosity profile from the DNS is far from constant. It has a clear maximum and a
  specific shape, which depends on Re.
  So, the authors took an empirical (engineering) approach on page 14 (line 403). They estimated a bulk
  eddy viscosity indirectly. This is a major weakness of this paper and could be avoided. Please see the
  following paper where we analytically derived the eddy viscosity profile from the Ekman equations.
  Since this is a relatively new paper, the authors might not have come across it. That is why I am revealing
  my identity.
  \begin{quotation}Basu, Sukanta, and Albert AM Holtslag. “A novel approach for deriving the stable
  boundary layer height and eddy viscosity profiles from the Ekman
  equations.” Boundary-Layer Meteorology 187.1 (2023): 105-115.
  \end{quotation}
  The authors might want to revise their derivations based on our analytical eddy viscosity profile. If that
  is difficult to do, I suggest highlighting the limitations of their current bulk approach. Also, emphasize
  the limitations of Eq. (10a) as it assumes constant eddy viscosity.}
  {We agree that eddy viscosity is not constant as a function of height. In fact, we even discuss the height dependency of eddy viscosity in Fig. 2c,d / Sec. 2. At the same time, the revised Ekman model / Eq. (10) makes use of Ekman's assumption of height-constant eddy viscosity. The important point here is that this is only used in the upper part of the boundary layer, where variation of the eddy viscosity is not the dominant mechanism for the wind profile. We find excellent agreement above $z^-\approx 0.5$ (cf. Fig. 6a/b in the manuscript). We are aware that deviations from the idealized Ekman profile below $z^-\approx 0.5$ are due to variations of eddy viscosity. 
  The key novelty in our approach regarding the Ekman spiral is, however, to apply it only in parts of the outer layer and to acknowledge its inappropriateness for most of what is commonly considered the turbulent boundary layer (about 2/3 of $\delta$). Across this range, where we argue to apply the Ekman model the eddy viscosity is not constant, but its variation appears limited, also in view of our own results presented in Sec. 2. We do hence not see an advantage to incorporate the model using variable eddy viscosity suggested by Basu et al. (2023) but rather understand this as an alternative approach, whichs is now mentioned in the revised manuscript.}{}
